\Large Advanced Simulation and Visualization of Fluids\\ in Computer Graphics \\ \large Divergence-Free Smoothed Particle Hydrodynamics

### 2016-12-13 report/paper.tex
gradient kernel in report
changed template for tex

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{balance}
\usepackage[nooneline,hang,it,IT]{subfigure}

% Our packages
\usepackage{listings} % To write code in latex
\usepackage{color}

@@

\lstset{style=mystyle}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{\LARGE TN1008 \\ Advanced Simulation and Visualization of Fluids in Computer Graphics \\ \Large Divergence-Free Smoothed Particle Hydrodynamics}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Ronja Grosz, rongr946\\ Isabell Jansson, isaja187\\ Jonathan Bosson, jonbo665}
\date{\today}
%\authorfooter{
%% insert punctuation at end of each item
%\item
 %Ronja Grosz is a student at Link\"oping University, Sweden, e-mail: rongr946@student.liu.se.
%}

%% Abstract section.
\abstract{ 
}
%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Divergence-free, SPH, divergence correction, density correction.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\firstsection{Introduction}
\maketitle 
    %explain the context of the work:
	%What exactly is the problem?
	%What have you created?
    Smoothed particle hydrodynamics, \textit{SPH}, is a method that was first implemented in 1977 for astrophysical simulations by Gingold et al.~\cite{firstSPH}.
    Since then SPH has become a popular method for complex water simulations.
    SPH is a mesh-free Lagrangian method where the particles move in space and change physical properties as time progresses.
    % More about Lagrangian and maybe navier stokes?

@@

\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
function performSimulation
    for all particles $i$ do //Init neighbourhoods
        find neighbourhoods $N_i(0)$
    for all particles $i$ do //Init $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$
        compute densities $\rho_i(0)$
        compute factors $\alpha_i(0)$
    while (t < $t_{max}$) do //Start simulation loop
        adapt time step $\Delta{t}$ 
        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
        for all particles $i$ do //Update positions
            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
        for all particles $i$ do //Update neighbourhoods
            find neighbourhoods $N_i(t + \Delta{t})$
        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
            compute densities $\rho_i(0)$
            compute factors $\alpha_i(0)$
        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$

\end{lstlisting}

@@

    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates for the cells and the fourth dimension is for storing the particles belonging to the cell.
    The amount of cells are decided by dividing the scene into cells of length $H$.
    The particles are then assigned to a cell according to equation~ref{eq:assignCell}.
    If the particle moves out of its cell it is then assigned to the new cell.

@@

\subsection{Divergence solver}
\subsection{Density solver}
\subsection{Kernel}
\subsection{Navier-stokes}
% Write about non pressure forces, velocity osv..?
\subsection{Adapted time step}
\subsection{Density and alpha factors}
\subsection{Screen space fluid rendering}

@@

%\bibliographystyle{abbrv}
\bibliographystyle{ieeetr}
\bibliography{./refs}

\end{document}
\grid

>>>>

\title{\Large Advanced Simulation and Visualization of Fluids\\ in Computer Graphics \\ \large Divergence-Free Smoothed Particle Hydrodynamics}

\author{\begin{tabular}{ccc}
    Isabell Jansson & Ronja Grosz  & Jonathan Bosson \\
    \small isaja187@student.liu.se &\small rongr946@student.liu.se &\small jonbo665@student.liu.se \\
\end{tabular}}

\date{\today}

\documentclass[12pt, twocolumn]{article}

% Our packages
\usepackage{graphicx}   % Images
\usepackage{amsmath}    % Math
\usepackage{amssymb}    % Math
\usepackage{listings}   % To write code in latex
\usepackage{color}      % Custom colors

@@

\lstset{style=mystyle}
% END our packages

\begin{document}
\twocolumn[
\begin{@twocolumnfalse}
\maketitle

\begin{abstract}
    This is the paper's abstract \ldots
    \\{\bf Index Terms:} Divergence-free, SPH, divergence correction, density correction.
\end{abstract}

\end{@twocolumnfalse}
]

\section{Introduction}
    %explain the context of the work:
	%What exactly is the problem?
	%What have you created?
    Smoothed particle hydrodynamics, \textit{SPH}, is a method that was first implemented in 1977 for astrophysical simulations by Gingold et al.~\cite{firstSPH}.
    Since then SPH has become a popular method for complex water simulations.
    SPH is a mesh-free Lagrangian method where the particles move in space and change physical properties as time progresses.
    % More about Lagrangian and maybe navier stokes?

@@

\begin{lstlisting}[language=C++, mathescape=true, escapeinside={(*}{*)}]
function performSimulation
    for all particles $i$ do //Init neighbourhoods
        find neighbourhoods $N_i(0)$
    for all particles $i$ do //Init $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$
        compute densities $\rho_i(0)$
        compute factors $\alpha_i(0)$
    while (t < $t_{max}$) do //Start simulation loop
        adapt time step $\Delta{t}$ 
        for all particles $i$ do //Predict velocities $\color{codegreen}\mathbf{v}_i^*$
            $\mathbf{v}_i = \mathbf{v}_i + \Delta{t}\mathbf{F}_i^{adv}/m_i$
        correctDensityError($\alpha, \mathbf{v}_i$) //Fulfill $\color{codegreen}\rho^* - \rho_0 = 0$
        for all particles $i$ do //Update positions
            $x_i(t + \Delta{t}) = x_i(t) + \Delta{t}\mathbf{v}_i^*$
        for all particles $i$ do //Update neighbourhoods
            find neighbourhoods $N_i(t + \Delta{t})$
        for all particles $i$ do //Update $\color{codegreen}\rho_i$ and $\color{codegreen}\alpha_i$ 
            compute densities $\rho_i(0)$
            compute factors $\alpha_i(0)$
        correctDivergenceError($\alpha, \mathbf{v}^*$) //Fulfill $\color{codegreen}\frac{D\rho}{Dt} = 0$
\end{lstlisting}

@@

    The cell list is implemented by using a four dimensional vector where the first three dimensions are for the x, y and z coordinates for the cells and the fourth dimension is for storing the particles belonging to the cell.
    The amount of cells are decided by dividing the scene into cells of length $H$.
    The particles are then assigned to a cell according to Equation ~\ref{eq:assignCell}.
    If the particle moves out of its cell it is then assigned to the new cell.

@@

% Should be mentioned early since it is used in mostly of the algorithms
\subsection{Kernel}
    A kernel function is used to simulate how particle-particle interactions decrease with the distance between the current particle and its neighbours. In SPH simulations this is an approximation of the Gaussian kernel function. Different kernels have been tested in previous works i.e. the poly6 kernel, the spiky kernel and the cubic spline kernel. According to Bender et. al.~\cite{bender} the cubic spline kernel presented by Monaghan~\cite{monaghan} was used. The kernel is described by Equation~\ref{eq:kernel}, where $q(x)=\frac{\left \| x \right \|}{h}$, $x$ is the distance between the current particle and a neighbour particle and $h$ is the support radius for the kernel. Particles further away than the support radius will not affect the current particle.

    \begin{equation} \label{eq:kernel}
        W_h(q(x)) =  \frac{1}{\pi h^3} \left\{\begin{matrix}
        1 - \frac{3}{2}q^2 + \frac{3}{4}q^3 & 0 \leqslant q < 1 \\ 
        \\
        \frac{1}{4}(2-q)^3 & 1 \leqslant q < 2\\ 
        \\
        0 & q \geqslant 2
      \end{matrix}\right.
    \end{equation}

    The algorithm does also require the kernel gradient. To reduce the computational effort and memory requirements Bender et. al.~\cite{bender} introduce a scalar function $g(q) = \frac{\partial W_h}{\partial q} \cdot \frac{1}{h \left \| x \right \|}$. The gradient kernel is then calculated by $\partial W_h(q(x)) = x \cdot g(x)$. The gradient kernel is described by Equation ~\ref{eq:gradKernel}.

    \begin{equation} \label{eq:gradKernel}
        \partial W_h(q(x)) =  x \cdot \frac{1}{h \left \| x \right \|} \cdot \frac{1}{\pi h^3} \left\{\begin{matrix}
        - 3q + \frac{9}{4}q^2 & 0 \leqslant q < 1 \\ 
        \\
        - \frac{3}{4}(2-q)^2 & 1 \leqslant q < 2\\ 
        \\
        0 & q \geqslant 2
      \end{matrix}\right.
    \end{equation}
    % maybe specify x here as well

    It is important to use the same kernel function for both $W_h$ and $\nabla W_h$ to get the prediction and the correction step to be compatible to each other.

\subsection{Divergence solver}
\subsection{Density solver}

\subsection{Navier-stokes}
% Write about non pressure forces, velocity osv..?
\subsection{Adapted time step}
\subsection{Density and alpha factors}
\subsection{Screen space fluid rendering}

@@

\bibliographystyle{ieeetr}
\bibliography{./refs}

\end{document}
